---
title: "GFK_Preprocessing"
author: "Orla Mallon"
date: "17/05/2022"
output: html_document
---

This script is the first in a series of 4 scripts which take messy panel data and convert it into a format which a hidden markov model can read. The series of scripts are as follows: 

1_GFK_Preprocessing.Rmd  = Initial merging and cleaning of the raw data files 
2_Categorising_Products = Finding and categorizing plant-based products 
3_Coding_Cateogries = Matching plant-based sales and coding these for the model
4_HMM_Model = Modelling the data using a hidden markov model 

The current script deals with the following steps: 

Part 1: Initial File Cleaning and Merging 
i) Loading the household, sales, and product data for 2018 - 2021
ii) Merging the old and new product data files 
iii) Cleaning the sales data 
iv) Finding the households which are activitely participating in the data 
v) Crop the sales data to include only active households 
vi) Merge the sales and product data

Part 2: Remove sales from categories of known confusion 
- Removes vegetables, plants, and oils from the sales data to be categorised 

**Output of Script: A clean dataframe of sales and product data merged ready to be categorised. **

## Setup 
```{r setup}
# -- working directory -- 
setwd("~/Desktop/dev/Data_Science/")
getwd()

# -- packages -- 
install.packages("readxl")
library("readxl") # for reading excel files 
library(tidyverse, dplyr) # data wrangling 
library(rio) # for reading stam files 
library(haven) # for reading stam files 

# -- load data -- 
# Sales for 2018, 2019, 2020, 2021 
sales_2018 <- read.csv("data/sales_data/Year_2018.csv") # (2,155,532 obs)
sales_2019 <- read.csv("data/sales_data/Year_2019.csv") # (2,096,983 obs)
sales_2020 <- read.csv("data/sales_data/Year_2020.csv", sep = ";") # (2,927,867 obs)
sales_2021 <- read.csv("data/sales_data/Year_2021.csv", colClasses="factor") # (2,212,234 obs)

# Household data 
weighted_HH <- read.csv("data/background_data/weighted_HH.csv") # 3898 obs 
households_2021 <- import("~/OneDrive - Aarhus Universitet/GFK Data/data_2021/data/stamgfk2022.sas7bdat")

# Product data 
articles <- read.csv("data/product_data/articles_2006_2021.csv", colClasses="factor")
articles_old <- read.csv("~/OneDrive - Aarhus Universitet/GFK Data/4. Shopping Data (Products)/articles2006-2020.csv", colClasses="factor")

# -- reclassify data columns -- 
articles$EanCodeText <- as.factor(articles$EanCodeText)
plant_based_data$`Bar Code` <- as.factor(plant_based_data$`Bar Code`)
```

### Merging old and new product data
The product data consists of information relating to every product's label, brand, and product group etc (see data descriptions). These files are used to contextualise the sales data and tell us more about the purchase made. 

The main column of interest is the eankode, which is how products are matched to the sales data. This is the product's 'barcode' or unique identifier number, which is given to the product when it is launched in the market. Sometimes products of a similar type and brand share eankode's, but this is rare. Thus we can be quite confident that the product files are able to give us an accurate idea of exactly what the consumer purchased. The eankode is the only link between the sales and product data. 

As the 2021 data has only recently been released, there are new products included in the product file. We also need to have the products from earlier years as we are working with data stretching back to 2018. Thus in the below section, the two product files will be merged in such a way that we can be sure to have all rows for both the old and new product information. 
```{r merge_product_files}
# -- STEP ONE: Make sure the 2 product (articles) dataframes are the same with no overlaps-- 

#1. Cut the private label columns from the old articles files 
articles_old <- articles_old %>% select(1:27)

#2. Get a a quick overview of the classes of each df column -- 
# articles 
table(sapply(articles, class)) # There are 25 character columns and 2 integer columns 

# old articles 
table(sapply(articles_old, class)) # There are 17 character columns and 10 integer

#3. Find the mismatched columns and convert them to factor columns (can handle alphanumeric)
sapply(articles, class)

# articles 
articles$EanCodeText <- as.factor(articles$EanCodeText)
articles$DType <- as.factor(articles$DType)
articles$FType <- as.factor(articles$FType)
articles$GType <- as.factor(articles$GType)
articles$IType <- as.factor(articles$IType)
articles$JType <- as.factor(articles$JType)
articles$KType <- as.factor(articles$KType)
articles$LType <- as.factor(articles$LType)
articles$MType <- as.factor(articles$MType)
articles$EanCodeNABS <- as.factor(articles$EanCodeNABS)

# articles old 
articles_old$EanCodeText <- as.factor(articles_old$EanCodeText)
articles_old$DType <- as.factor(articles_old$DType)
articles_old$FType <- as.factor(articles_old$DType)
articles_old$GType <- as.factor(articles_old$DType)
articles_old$IType <- as.factor(articles_old$DType)
articles_old$JType <- as.factor(articles_old$DType)
articles_old$KType <- as.factor(articles_old$DType)
articles_old$LType <- as.factor(articles_old$DType)
articles_old$MType <- as.factor(articles_old$DType)
articles_old$EanCodeNABS <- as.factor(articles_old$EanCodeNABS)

#4. Find the eankodes which are in both dataframes
# This ensures we will not have overlaps or conflicting information 
ean_in_both <- semi_join(articles_old, articles, by = "EanCodeText") # 64869 
# remove these from old articles 
articles_old <- anti_join(articles_old, ean_in_both) # 395,827 obs 
rm(ean_in_both)


# -- STEP TWO: Ensure each data frame has unique rows (i.e., no duplicate rows) -- 
## articles 
articles_distinct <- articles %>% distinct(EanCodeText, 
                                           ProductGroupCode, 
                                           ArticleText, 
                                           .keep_all = TRUE ) # 242,477 obs 

# check the rows which are removed 
articles_removed <- anti_join(articles, articles_distinct) # 36,307 
articles_removed <- articles_removed %>% group_by(EanCodeText) %>% mutate(ean_count = n())

# reduce articles to include only one count of each Eancode (after checking)
articles <- articles %>% distinct(EanCodeText, 
                                  .keep_all = TRUE) # 42216 obs 

## articles_old 
articles_old_distinct <- articles_old %>% distinct(EanCodeText, 
                                                   ProductGroupCode, 
                                                   ArticleText, 
                                                   .keep_all = TRUE) # 389,449 obs 

# check the removed rows 
articles_old_removed <- anti_join(articles_old, articles_old_distinct)
articles_old_removed <- articles_old_removed %>% 
  group_by(EanCodeText) %>% 
  mutate(ean_count = n())

# reduce articles_old to include only one count of each eankode 
articles_old <- articles_old %>% distinct(EanCodeText, 
                                          .keep_all = TRUE) # 144,628 obs 

# -- STEP THREE: Combine the 2 articles files -- 
# Combine the 2 articles files into one 
articles_new <- rbind(articles, articles_old) # 186844 obs 
articles_new <- articles_new %>% group_by(EanCodeText) %>% mutate(count = n()) # check for dups
articles <- articles_new # make this the new articles dataframe 

# clean up environment 
rm(articles_distinct, articles_old_distinct, 
   articles_old_removed, articles_removed, 
   articles_new, articles_old)

```

Additional notes: After thoroughly exploring the product file, the decision was taken to reduce this down to include only one entry for each eankode. This was because the eankode was often capturing the same type of product with slight difference between such as flavour etc, which causes problems when we merge it with the sales data. Other eankodes were problematic as they spanned many different categories and it was though to not be useful to overcomplicate the data by including them, thus they've been removed so that the product file has only one entry (row) for each eankode. 

### Clean up the sales data 
The sales is the records of every purchase made by households across the year. These contain millions of rows of data, and are often messy and in need of some cleaning. The steps taken to handle these files are outlined below: 
```{r clean_sales_data}
# -- Re-classify the ean kode columns -- 
sales_2017$eankode <- as.factor(sales_2017$eankode)
sales_2018$eankode <- as.factor(sales_2018$eankode)
sales_2019$eankode <- as.factor(sales_2019$eankode)
sales_2020$EAN <- as.factor(sales_2020$EAN)
sales_2021$eankode <- as.factor(sales_2021$eankode)

# -- Clip data to include only columns of interest -- 
sales_2017 <- sales_2017 %>% select(hhnr, eankode, year, sasdato, shop, tid, ANTAL)
sales_2018 <- sales_2018 %>% select(hhnr, eankode, year, sasdato, shop, tid, ANTAL)
sales_2019 <- sales_2019 %>% select(hhnr, eankode, year, sasdato, shop, tid, ANTAL)
sales_2020 <- sales_2020 %>% select(HHNR, EAN, YEAR, DAY, MONTH, SHOP, TID, ANTAL)
sales_2021 <- sales_2021 %>% select(hhnr, eankode, year, sasdato, shop, tid, ANTAL)

# -- rename sales_2020 columns -- 
sales_2020 <- sales_2020 %>% 
  rename(
    hhnr = HHNR, 
    eankode = EAN, 
    year = YEAR, 
    day = DAY, 
    month = MONTH, 
    shop = SHOP, 
    tid = TID)

# -- rename date column for 2017, 2018, and 2019 -- 
sales_2017 <- sales_2017 %>% rename(date = sasdato)
sales_2018 <- sales_2018 %>% rename(date = sasdato)
sales_2019 <- sales_2019 %>% rename(date = sasdato)
sales_2021 <- sales_2021 %>% rename(date = sasdato)

# - reclassify date column into 'Date' column -- 
sales_2017$date <- as.Date(sales_2017$date)
sales_2018$date <- as.Date(sales_2018$date)
sales_2019$date <- as.Date(sales_2019$date)
sales_2021$date <- as.Date(sales_2021$date)
class(sales_2019$date)

# -- Create date column from day month year (style = YYYY-MM-DD) in sales_2020 -- 
sales_2020$date <- as.Date(with(sales_2020, paste(year, month, day,sep="-")), "%Y-%m-%d")
class(sales_2020$date)

# reorder and delete redundant columns 
sales_2020 <- sales_2020 %>% select(hhnr, eankode, year, date, shop, tid, ANTAL)
```

### Keep sales data from active households 
The weighted_HH dataset contains information on whether the household has actively participated in the panel data for each year. A 1 indicates they have actively participated (i.e., been recording sales), while a 0 indicates the household has not been active.

The 2021 data has not yet been added to the weighted household dataset so we'll first need to record which households have been actively participating in the panel data. This is defined as households with at least 100 sales records throughout 2021. 
```{r find_active_households}
# -- Clean up weighted_HH df -- 
# delete redundent columns 
weighted_HH <- weighted_HH %>% select(X, X2017, X2018, X2019, X2020)  

# delete first row 
weighted_HH <- weighted_HH[-c(1),] # 3897 obs

# rename rows 
weighted_HH <- weighted_HH %>% rename(
  hh_id = X,
  year_2017 = X2017, 
  year_2018 = X2018,
  year_2019 = X2019, 
  year_2020 = X2020) # (3897 obs)

# -- Create a column for the 2021 households -- 
# get a list of the active 2021 households 
active_households_2021 <- sales_2021 %>% group_by(hhnr) %>% mutate(purchase = n())
active_households_2021 <- active_households_2021 %>% distinct(hhnr, .keep_all = T) # 3086 

# check how many households from 2021 are included in weighted_HH 
hh_check <-weighted_HH[weighted_HH$hh_id %in% active_households_2021$hhnr, ] 
hh_check <- hh_check %>% distinct(hh_id) # 1969 households 

# crop 2021 households to include only the ones with more than 100 purchases across the year 
active_households_2021 <- active_households_2021 %>% subset(purchase > 99) # 2475 households 
active_households_2021 <- active_households_2021 %>% select(hhnr)

# check how many active 2021 households are in the hh_check 
hh_check <- weighted_HH[weighted_HH$hh_id %in% active_households_2021$hhnr, ] # 1887 households
hh_check$year_2021 <- as.character("1")
hh_check <- hh_check %>% select(hh_id, year_2021)

# merge the 2021 households to weighted_HH 
weighted_HH <- merge(weighted_HH, hh_check, 
              by = "hh_id",
              all.x = T, all.y = F ) #3897 obs 

# make the NA's into 0s 
weighted_HH$year_2021[is.na(weighted_HH$year_2021)] <- "0"

# Sum the 5 years and if the total = 5, it means they have been actively participating. 
# reclassify the columns into integers 
weighted_HH$year_2017 <- as.integer(weighted_HH$year_2017)
weighted_HH$year_2018 <- as.integer(weighted_HH$year_2018)
weighted_HH$year_2019 <- as.integer(weighted_HH$year_2019)
weighted_HH$year_2020 <- as.integer(weighted_HH$year_2020)
weighted_HH$year_2021 <- as.integer(weighted_HH$year_2021)

# sum the 4 years 
weighted_HH$participation_score <- weighted_HH$year_2017 + weighted_HH$year_2018 + 
  weighted_HH$year_2019 + weighted_HH$year_2020 + weighted_HH$year_2021

# remove households with participation_score of less than 4 
weighted_HH <- weighted_HH %>% subset(participation_score == 5) # 1167 households 

# reclassify hh_id 
weighted_HH$hh_id <- as.factor(weighted_HH$hh_id)

# Save household data csv 
write.csv(weighted_HH, "data/background_data/households_updated.csv")

# clean up the environment 
rm(active_households_2021, hh_check)
```

### Cut sales data to only include active households 
Having now identified the households which are active, and thus of interest to the study, we will crop away any data from households which are not in this list. These may be households who sporadically participate or whom have dropped out of the study. The steps taken are outlined below: 
```{r keep_active_household_data}
# -- reclassify the hhnr column -- 
sales_2017$hhnr <- as.factor(sales_2017$hhnr)
sales_2018$hhnr <- as.factor(sales_2018$hhnr)
sales_2019$hhnr <- as.factor(sales_2019$hhnr)
sales_2020$hhnr <- as.factor(sales_2020$hhnr)
sales_2021$hhnr <- as.factor(sales_2021$hhnr)
sum(is.na(sales_2021$hhnr)) # check there are no empty hh entries

# Match the sales with household numbers in the weighted_hh file 
sales_2017 <- sales_2017[sales_2017$hhnr %in% weighted_HH$hh_id, ] # (1283,802 obs)
sales_2018 <- sales_2018[sales_2018$hhnr %in% weighted_HH$hh_id, ] # (1300,648 obs)
sales_2019 <- sales_2019[sales_2019$hhnr %in% weighted_HH$hh_id, ] # (1233,838 obs)
sales_2020 <- sales_2020[sales_2020$hhnr %in% weighted_HH$hh_id, ] # (1587,493 obs)
sales_2021 <- sales_2021[sales_2021$hhnr %in% weighted_HH$hh_id, ] # (1213,765 obs)

# Combine all sales into one 
sales_data <- rbind(sales_2017, sales_2018, sales_2019, sales_2020, sales_2021) #(6619,546 obs)
print(sum(is.na(sales_data$hhnr)))
print(sum(is.na(sales_data$eankode)))

# Save sales data 
write.csv(sales_data, "data/sales_data/all_sales_data.csv", row.names = FALSE)

# clean up environment 
rm(sales_2017, sales_2018, sales_2019, sales_2020, sales_2021)
```


#### Merge sales and product data 
With the sales and product data cleaned and cropped, we can now merge the files so that we have all the product information relevant to the sale. These two files are merged using the eankode column: 
```{r merge_sales_and_product}
# -- Ensure sales data and product data ean codes are factors -- 
class(sales_data$eankode)
class(articles$EanCodeText)
#sales_data$eankode <- as.factor(sales_data$eankode)
#articles$EanCodeText <- as.factor(articles$EanCodeText)

# -- Clip articles file (product data) -- 
articles <- articles %>% select(EanCodeText, ProductGroupCode, PgText, ArticleText, Cmrk, ManufactorName, BrandText) # 186844 obs 

# -- Merge the files by ean -- 
sales_data <- merge(sales_data, articles, 
                    by.x = "eankode", by.y = "EanCodeText",
                    all = TRUE) # 6745359 obs

# -- Clean up the merged data -- 
# 1. Remove rows where no eancode has been recorded 
sum(sales_data$eankode == 0) # 456,493 sales with no eancode recorded 
sales_data <- sales_data[!(sales_data$eankode== 0),] # (6288,866 obs)

# 2. Remove rows where no household has been recorded 
sum(is.na(sales_data$hhnr)) # 125813 sales with no household recorded 
sales_data <- sales_data[!(is.na(sales_data$hhnr)),] # 6163053 obs

# 3. Remove rows where no ArticleText has been recorded (i.e., does not know about product)
sum(is.na(sales_data$ArticleText)) # 808,357 
sales_data <- sales_data[!(is.na(sales_data$ArticleText)),] #5354,696 obs 

# 3. Remove duplicates 
sales_data <- sales_data %>% distinct() # 5308,638 obs 

# -- Quality checks: ensure there are no NAs in important column -- 
print(sum(is.na(sales_data$eankode)))
print(sum(is.na(sales_data$hhnr)))
print(sum(is.na(sales_data$ArticleText)))
print(sum(is.na(sales_data$date)))

# -- Save dataframe -- 
write.csv(sales_data, "data/sales_data/sales_with_product_data.csv", row.names = FALSE)

```

## Part 2: Exclude irrelevant categories of Known confusion
The below rows are removed in advance before exploring as they are known to cause confusion within the pulling of data. This allows better filtering by searching for words such as 'plant-base' without catching plants etc. 

Vegetables = removed because they refer to base fruit and vegetable products, which are not a plant-based alternative but can be caught in subsetting. 

Plants = removed because they often have the word "plant" in their description, but are referring to flowers, seeds, and plant pots. 

Oils = removed as sometimes recorded as vegan products but they are typically plant-based and therefore not an alternative

```{r filter_product_groups}
# -- reload sales data if not running script from top -- 
#sales_reloaded <- read.csv("data/sales_data/sales_with_product_data.csv")

# -- Vegetables (Grøntsager (047)) -- 
# First check they have no articles of interest 
vegetables <- sales_data %>% subset(ProductGroupCode == 47) # 565,477 obs  
# get only distinct article texts and eyeball to ensure all should be removed 
vegetables_unique <- vegetables %>% distinct(ArticleText, .keep_all = T) # 1076
# remove from the sales data 
sales_data <- anti_join(sales_data, vegetables) # (sales =4743,161)
# remove dfs 
rm(vegetables, vegetables_unique)

# -- Plants (Blomster (080)) -- 
plants <- sales_data %>% subset(ProductGroupCode == 80) #28,917 obs 
plants_unique <- plants %>% distinct(ArticleText, .keep_all = T) # check
sales_data <- anti_join(sales_data, plants) # remove from sales (sales = 4714,244 obs)
# also remove plant pots 
plant_pots <- sales_data[grep("Potteplante", sales_data$ArticleText),] # 6779 obs
sales_data <- anti_join(sales_data, plant_pots) # 4707,465 obs
# also remove random plant product 
plant_1stk <- sales_data[grep("Plante 1 stk", sales_data$ArticleText),]	# 98 obs 
sales_data <- anti_join(sales_data, plant_1stk) # (sales = 4707367 obs)
rm(plants, plants_unique, plant_pots, plant_1stk)

# -- Oils (Mad og Spiseolie (050)) -- 
oils <- sales_data %>% subset(ProductGroupCode == 50) # 13036 
oils_unique <- oils %>% distinct(ArticleText, .keep_all = T)
sales_data <- anti_join(sales_data, oils) # (sales = 4694331 obs)
rm(oils, oils_unique)

# -- Re-save dataframe -- 
write.csv(sales_data, "data/sales_data/sales_with_product_data.csv", row.names = FALSE)

```

This is the end 