---
title: "Coding Categories"
author: "Orla Mallon"
date: "26/05/2022"
output: html_document
---

This script takes the plant-based data files and codes them into a way the model is able to understand.

## Setup 
```{r setup}
# -- working dircetory -- 
setwd("~/Desktop/dev/Data_Science/")
getwd()

# -- dependencies -- 
library(tidyverse, dplyr)
library(ggplot2)

# -- Load Data -- 
pb_dairy <- read.csv("data/categorized_sales/pb_dairy.csv", colClasses="factor")
pb_meat <- read.csv("data/categorized_sales/pb_meat.csv", colClasses="factor")
pb_eggs <- read.csv("data/categorized_sales/pb_eggs.csv", colClasses="factor")
pb_seafood <- read.csv("data/categorized_sales/pb_seafood.csv", colClasses="factor")
pb_readymeals <- read.csv("data/categorized_sales/pb_readymeals.csv", colClasses="factor")
pb_confectionery <- read.csv("data/categorized_sales/pb_confectionery.csv", colClasses="factor")
```

#### Fuctions 
```{r functions}
# -- Function to count purchases by household in a dataframe -- 
count_purchases <- function(df) { 
  counted_data <- df %>% group_by(hhnr) %>% 
    mutate(purchases = n()) %>% 
    distinct(hhnr, .keep_all = T) %>% 
    select(hhnr, purchases)
  return(counted_data)
}

# -- Function to rename columns for saving-- 
# rename columns 
rename_columns <- function(df){
  new_df <- df %>% 
    rename(
      hhnr = 1,
      Period_1 = 2,
      Period_2 = 3, 
      Period_3 = 4,
      Period_4 = 5,
      Period_5 = 6,
      Period_6 = 7,
      Period_7 = 8,
      Period_8 = 9)
  return(new_df)
}
```

#### Pre-Visualisation of the Data:
```{r viz}
p1_all_years <- categorised_data %>% group_by(year) %>% 
  ggplot(aes(x = year, fill = year)) + 
  geom_bar() + 
  theme_classic() + 
  labs(title = "Sales of Plant-Based Products by Year", 
       x = "Year", y = "") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14)) +
  theme(legend.position="none") + 
  scale_fill_brewer(palette="Spectral")
p1_all_years

p2_by_category <- pb_dairy %>% group_by(year) %>% 
  ggplot(aes(x = year, fill = year)) + 
  geom_bar() + 
  theme_classic() + 
  labs(title = "Sales of Dairy Products by Year", 
       x = "Year", y = "") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14)) +
  theme(legend.position="none") + 
  scale_fill_brewer(palette="Spectral")
p2_by_category

p3_by_category <- pb_meat %>% group_by(year) %>% 
  ggplot(aes(x = year, fill = year)) + 
  geom_bar() + 
  theme_classic() + 
  labs(title = "Sales of Meat Products by Year", 
       x = "Year", y = "") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14)) +
  theme(legend.position="none") + 
  scale_fill_brewer(palette="Spectral")
p3_by_category

p4_by_category <- pb_readymeals %>% group_by(year) %>% 
  ggplot(aes(x = year, fill = year)) + 
  geom_bar() + 
  theme_classic() + 
  labs(title = "Sales of Ready Meal Products by Year", 
       x = "Year", y = "") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14)) +
  theme(legend.position="none") + 
  scale_fill_brewer(palette="Spectral")
p4_by_category

p5_by_category <- pb_confectionery %>% group_by(year) %>% 
  ggplot(aes(x = year, fill = year)) + 
  geom_bar() + 
  theme_classic() + 
  labs(title = "Sales of Confectionery Products by Year", 
       x = "Year", y = "") + 
  theme(plot.title = element_text(hjust = 0.5, size = 14)) +
  theme(legend.position="none") + 
  scale_fill_brewer(palette="Spectral")
p5_by_category

```

Due to the low number of sales in the 'pb_eggs' (262 obs) and 'pb_seafood' (31 obs), these two categories will be dropped from the rest of the analysis. This means we will be looking for purchases per household in the 'Dairy', 'Meat', 'Ready Meals' and 'Confectionery' columns. 

In the below chunks,  the following steps are taken to code each household as either a purchase (1) or non purchase (0) for each category in each time step. 

i)The category's sales files are cut into 6 month chuncks (Jan 18 - Dec 21)
ii) The number of purchases for each household in each time chunk is calculated.
iii) Append sales data onto a list of the household IDs (A non sale becomes an NA)
iv) NA's are converted to 0's, and anything that is not 0 is converted to a 1 to reflect the household made a purchase. The volume of sales is disregarded. 
v) Each category is checked to see how many households have purchased products in this category across the 4 years. 



#### Dairy 
```{r dairy_time_periods}
# -- Step 1: Cut the data into 6 month chunks -- 

# Make the Date column into a date column 
pb_dairy$date <- as.Date(pb_dairy$date)

# -- Time Periods -- 
# -- 2018 -- 
P1_dairy <- pb_dairy[pb_dairy$date >= "2018-01-01" & pb_dairy$date <= "2019-06-30",] # 7335 
P2_dairy <- pb_dairy[pb_dairy$date >= "2018-07-01" & pb_dairy$date <= "2019-12-31",] # 6443

# -- 2019 -- 
P3_dairy <- pb_dairy[pb_dairy$date >= "2019-01-01" & pb_dairy$date <= "2019-06-30",] # 2273 
P4_dairy <- pb_dairy[pb_dairy$date >= "2019-07-01" & pb_dairy$date <= "2019-12-31",] # 1637 

# -- 2020 -- 
P5_dairy <- pb_dairy[pb_dairy$date >= "2020-01-01" & pb_dairy$date <= "2020-06-30",] # 1837 
P6_dairy <- pb_dairy[pb_dairy$date >= "2020-07-01" & pb_dairy$date <= "2020-12-31",] # 1797

# -- 2021 -- 
P7_dairy <- pb_dairy[pb_dairy$date >= "2021-01-01" & pb_dairy$date <= "2021-06-30",] # 1696 
P8_dairy <- pb_dairy[pb_dairy$date >= "2021-07-01" & pb_dairy$date <= "2021-12-31",] # 1510 

```

```{r count_purchases}
# -- Step 2: Count the pb_dairy purchases per household -- 
P1_dairy <- count_purchases(P1_dairy) # 887 
P2_dairy <- count_purchases(P2_dairy) # 848 
P3_dairy <- count_purchases(P3_dairy) # 617 
P4_dairy <- count_purchases(P4_dairy) # 374 
P5_dairy <- count_purchases(P5_dairy) # 396
P6_dairy <- count_purchases(P6_dairy) # 394
P7_dairy <- count_purchases(P7_dairy) # 364
P8_dairy <- count_purchases(P8_dairy) # 351 
```

```{r code_purchases}
# -- Step 3: Match the sales data to the household -- 
# Create a data frame to append the data into --  
pb_households <- screened_households %>% distinct(hhnr)

# --2018 -- 
pb_households$Dairy_P1 <- P1_dairy$purchases[match(pb_households$hhnr,P1_dairy$hhnr)]
pb_households$Dairy_P2 <- P2_dairy$purchases[match(pb_households$hhnr,P2_dairy$hhnr)]

# -- 2019 -- 
pb_households$Dairy_P3 <- P3_dairy$purchases[match(pb_households$hhnr,P3_dairy$hhnr)]
pb_households$Dairy_P4 <- P4_dairy$purchases[match(pb_households$hhnr,P4_dairy$hhnr)]

# -- 2020 -- 
pb_households$Dairy_P5 <- P5_dairy$purchases[match(pb_households$hhnr,P5_dairy$hhnr)]
pb_households$Dairy_P6 <- P6_dairy$purchases[match(pb_households$hhnr,P6_dairy$hhnr)]

# --2021 -- 
pb_households$Dairy_P7 <- P7_dairy$purchases[match(pb_households$hhnr,P7_dairy$hhnr)]
pb_households$Dairy_P8 <- P7_dairy$purchases[match(pb_households$hhnr,P8_dairy$hhnr)]

# -- Step 4: Convert NAs to 0 and everything else to 1 -- 
# ensure NAs are 0s (this means no sale made)
sum(is.na(pb_households)) # 5041
pb_households[is.na(pb_households)] <- 0 

# split the house numbers away 
pb_household_split <- select(pb_households, -hhnr)
hhnrs <- pb_households %>% select(hhnr)

# set everything that isn't a 0 to 1 (meaning purchase made)
pb_household_split[pb_household_split != 0] <- 1

# bind the columns back together 
pb_households <- cbind(hhnrs, pb_household_split)
```

```{r find_hhnrs_exclude}
# -- Step 5: Check number of households who have made pb_dairy purchases -- 
# -- Find households who made no pb_dairy purchases and exclude them from data -- pb_households <- pb_households %>% 
  mutate(Total = select(., Dairy_P1:Dairy_P8) %>% rowSums(na.rm = TRUE))

# -- Exclude households who made no purchases -- 
excluded_dairy <- pb_households %>% subset(Total < 1) # 170 households 

# Thus there are 989 active households for the dairy category. 
```

```{r clean_dairy}
# clean up environment
rm(P1_dairy, P2_dairy, P3_dairy, P4_dairy, P5_dairy, P6_dairy, P7_dairy, P8_dairy)

```


#### Meat 
```{r meat_time_periods}
# -- Step 1: Cut the data into 6 month chunks -- 
# Make the Date column into a date column 
pb_meat$date <- as.Date(pb_meat$date)

# -- Time Periods --
# -- 2018 -- 
P1_meat <- pb_meat[pb_meat$date >= "2018-01-01" & pb_meat$date <= "2018-06-30",] # 445
P2_meat <- pb_meat[pb_meat$date >= "2018-07-01" & pb_meat$date <= "2018-12-31",] # 313

# -- 2019 -- 
P3_meat <- pb_meat[pb_meat$date >= "2019-01-01" & pb_meat$date <= "2019-06-30",] # 412
P4_meat <- pb_meat[pb_meat$date >= "2019-07-01" & pb_meat$date <= "2019-12-31",] # 358  
 
# -- 2020 -- 
P5_meat <- pb_meat[pb_meat$date >= "2020-01-01" & pb_meat$date <= "2020-06-30",] # 321 
P6_meat <- pb_meat[pb_meat$date >= "2020-07-01" & pb_meat$date <= "2020-12-31",] # 279 

# -- 2021 -- 
P7_meat <- pb_meat[pb_meat$date >= "2021-01-01" & pb_meat$date <= "2021-06-30",] # 276 
P8_meat <- pb_meat[pb_meat$date >= "2021-07-01" & pb_meat$date <= "2021-12-31",] # 231 

```

```{r count_purchases}
# -- Step 2: Count the pb_meat purchases per household -- 
P1_meat <- count_purchases(P1_meat) #184  
P2_meat <- count_purchases(P2_meat) #147
P3_meat <- count_purchases(P3_meat) #171
P4_meat <- count_purchases(P4_meat) #148
P5_meat <- count_purchases(P5_meat) #145
P6_meat <- count_purchases(P6_meat) #136
P7_meat <- count_purchases(P7_meat) #114
P8_meat <- count_purchases(P8_meat) #104
```

```{r code_purchases}
# -- Step 3: Match the sales data to the household -- 
# Create dataframe to append data into 
pbm_households <- screened_households %>% distinct(hhnr)

# -- 2018 -- 
pbm_households$Meat_P1 <- P1_meat$purchases[match(pbm_households$hhnr,P1_meat$hhnr)]
pbm_households$Meat_P2 <- P2_meat$purchases[match(pbm_households$hhnr,P2_meat$hhnr)]

# -- 2019 -- 
pbm_households$Meat_P3 <- P3_meat$purchases[match(pbm_households$hhnr,P3_meat$hhnr)]
pbm_households$Meat_P4 <- P4_meat$purchases[match(pbm_households$hhnr,P4_meat$hhnr)]

# -- 2020 -- 
pbm_households$Meat_P5 <- P5_meat$purchases[match(pbm_households$hhnr,P5_meat$hhnr)]
pbm_households$Meat_P6 <- P6_meat$purchases[match(pbm_households$hhnr,P6_meat$hhnr)]

# -- 2021 -- 
pbm_households$Meat_P7 <- P7_meat$purchases[match(pbm_households$hhnr,P7_meat$hhnr)]
pbm_households$Meat_P8 <- P8_meat$purchases[match(pbm_households$hhnr,P8_meat$hhnr)]

# -- Code the purchases -- 
# ensure NAs are 0s (this means no sale made)
sum(is.na(pbm_households)) #8123 
pbm_households[is.na(pbm_households)] <- 0 

# split the house numbers away 
pbm_household_split <- select(pbm_households, -hhnr)
hhnrs_pbm <- pbm_households %>% select(hhnr)

# set everything that isn't a 0 to 1 (meaning purchase made)
pbm_household_split[pbm_household_split != 0] <- 1

# bind the columns back together 
pbm_households <- cbind(hhnrs_pbm, pbm_household_split)

```

```{r find_hhnrs_exclude}
# -- Step 5: Check number of households who have made pb_meat purchases -- 
# -- Find households who made no pb_meat purchases and exclude them from data -- 
pbm_households <- pbm_households %>% 
  mutate(Total = select(., Meat_P1:Meat_P8) %>% rowSums(na.rm = TRUE)) 

# -- Exclude households who made no purchases -- 
excluded_meat <- pbm_households %>% subset(Total < 1) # 673 households 

# This leaves 673 households which meet the 'Meat category' 
```

```{r clean_meat}
# clean up environment
rm(P1_meat, P2_meat, P3_meat, P4_meat, P5_meat, P6_meat, P7_meat, P8_meat)
```

#### Ready Meals 

```{r readymeal_time_periods}
# -- Step 1: Cut the data into 6 month chunks -- 
# Make the Date column into a date column 
pb_readymeals$date <- as.Date(pb_readymeals$date)

# -- Time Periods -- 
# -- 2018 -- 
P1_readymeals <- pb_readymeals[pb_readymeals$date >= "2018-01-01" & 
                               pb_readymeals$date <= "2018-06-30",] # 356
P2_readymeals <- pb_readymeals[pb_readymeals$date >= "2018-07-01" & 
                               pb_readymeals$date <= "2018-12-31",] # 299  

# -- 2019 -- 
P3_readymeals <- pb_readymeals[pb_readymeals$date >= "2019-01-01" & 
                               pb_readymeals$date <= "2019-06-30",] # 307  
P4_readymeals <- pb_readymeals[pb_readymeals$date >= "2019-07-01" & 
                               pb_readymeals$date <= "2019-12-31",] # 268 

# -- 2020 -- 
P5_readymeals <- pb_readymeals[pb_readymeals$date >= "2020-01-01" & 
                               pb_readymeals$date <= "2020-06-30",] # 281 
P6_readymeals <- pb_readymeals[pb_readymeals$date >= "2020-07-01" & 
                               pb_readymeals$date <= "2020-12-31",] # 194 

# -- 2021 -- 
P7_readymeals <- pb_readymeals[pb_readymeals$date >= "2021-01-01" & 
                               pb_readymeals$date <= "2021-06-30",] # 164 
P8_readymeals <- pb_readymeals[pb_readymeals$date >= "2021-07-01" & 
                               pb_readymeals$date <= "2021-12-31",] # 102 

```

```{r count_purchases}
# -- Step 2: Count the pb_readymeals purchases per household -- 
P1_readymeals <- count_purchases(P1_readymeals) #174  
P2_readymeals <- count_purchases(P2_readymeals) #164
P3_readymeals <- count_purchases(P3_readymeals) #183
P4_readymeals <- count_purchases(P4_readymeals) #162
P5_readymeals <- count_purchases(P5_readymeals) #145
P6_readymeals <- count_purchases(P6_readymeals) #109
P7_readymeals <- count_purchases(P7_readymeals) #95
P8_readymeals <- count_purchases(P8_readymeals) #67
```

```{r code_purchases}
# -- Step 3: Match the sales data to the household -- 
# Create dataframe to append data into 
pbr_households <- screened_households %>% distinct(hhnr)

# -- 2018 -- 
pbr_households$ReadyMeal_P1 <- P1_readymeals$purchases[match(pbr_households$hhnr,
                                                             P1_readymeals$hhnr)]
pbr_households$ReadyMeal_P2 <- P2_readymeals$purchases[match(pbr_households$hhnr,
                                                             P2_readymeals$hhnr)]

# -- 2019 -- 
pbr_households$ReadyMeal_P3 <- P3_readymeals$purchases[match(pbr_households$hhnr,
                                                             P3_readymeals$hhnr)]
pbr_households$ReadyMeal_P4 <- P4_readymeals$purchases[match(pbr_households$hhnr,
                                                             P4_readymeals$hhnr)]
# -- 2020 -- 
pbr_households$ReadyMeal_P5 <- P5_readymeals$purchases[match(pbr_households$hhnr,
                                                             P5_readymeals$hhnr)]
pbr_households$ReadyMeal_P6 <- P6_readymeals$purchases[match(pbr_households$hhnr,
                                                             P6_readymeals$hhnr)]

# -- 2021 -- 
pbr_households$ReadyMeal_P7 <- P7_readymeals$purchases[match(pbr_households$hhnr,
                                                             P7_readymeals$hhnr)]
pbr_households$ReadyMeal_P8 <- P8_readymeals$purchases[match(pbr_households$hhnr,
                                                             P8_readymeals$hhnr)]

# -- Code the purchases -- 
# ensure NAs are 0s (this means no sale made)
sum(is.na(pbr_households)) #8173 
pbr_households[is.na(pbr_households)] <- 0 

# split the house numbers away 
pbr_household_split <- select(pbr_households, -hhnr)
hhnrs_pbr <- pbr_households %>% select(hhnr)

# set everything that isn't a 0 to 1 (meaning purchase made)
pbr_household_split[pbr_household_split != 0] <- 1

# bind the columns back together 
pbr_households <- cbind(hhnrs_pbr, pbr_household_split)

```

```{r find_hhnrs_exclude}
# -- Step 5: Check number of households who have made pb_readymeals purchases -- 
# -- Find households who made no pb_meat purchases and exclude them from data -- 
pbr_households <- pbr_households %>% 
  mutate(Total = select(., ReadyMeal_P1:ReadyMeal_P8) %>% rowSums(na.rm = TRUE)) 

# -- Exclude households who made no purchases -- 
excluded_readymeal <- pbr_households %>% subset(Total < 1) # 682 households 

# Thus 477 households meet the readymeal criteria 
```

```{r clean_readymeals}
rm(P1_readymeals, P2_readymeals, P3_readymeals, P4_readymeals, P5_readymeals, P6_readymeals,
   P7_readymeals, P8_readymeals)
```

#### Confectionery 

```{r confectionery_time_periods}
# -- Step 1: Cut the data into 6 month chunks -- 
# Make the Date column into a date column 
pb_confectionery$date <- as.Date(pb_confectionery$date)

# -- Time Periods -- 
# -- 2018 -- 
P1_confectionery <- pb_confectionery[pb_confectionery$date >= "2018-01-01" & 
                                     pb_confectionery$date <= "2018-06-30",] # 1831 
P2_confectionery <- pb_confectionery[pb_confectionery$date >= "2018-07-01" & 
                                     pb_confectionery$date <= "2018-12-31",] # 1685  

# -- 2019 -- 
P3_confectionery <- pb_confectionery[pb_confectionery$date >= "2019-01-01" & 
                                     pb_confectionery$date <= "2019-06-30",] # 1794  
P4_confectionery <- pb_confectionery[pb_confectionery$date >= "2019-07-01" & 
                                     pb_confectionery$date <= "2019-12-31",] # 1563

# -- 2020 -- 
P5_confectionery<- pb_confectionery[pb_confectionery$date >= "2020-01-01" & 
                                    pb_confectionery$date <= "2020-06-30",] # 1822 
P6_confectionery <- pb_confectionery[pb_confectionery$date >= "2020-07-01" & 
                                     pb_confectionery$date <= "2020-12-31",] # 1530

# -- 2021 -- 
P7_confectionery <- pb_confectionery[pb_confectionery$date >= "2021-01-01" & 
                                     pb_confectionery$date <= "2021-06-30",] # 1169  
P8_confectionery <- pb_confectionery[pb_confectionery$date >= "2021-07-01" & 
                                     pb_confectionery$date <= "2021-12-31",] # 890 
```

```{r count_purchases}
# -- Step 2: Count the pb_confectionery purchases per household -- 
P1_confectionery <- count_purchases(P1_confectionery) #605  
P2_confectionery <- count_purchases(P2_confectionery) #578
P3_confectionery <- count_purchases(P3_confectionery) #547
P4_confectionery <- count_purchases(P4_confectionery) #524
P5_confectionery <- count_purchases(P5_confectionery) #584
P6_confectionery <- count_purchases(P6_confectionery) #530
P7_confectionery <- count_purchases(P7_confectionery) #404
P8_confectionery <- count_purchases(P8_confectionery) #359
```

```{r code purchases}
# -- Step 3: Match the sales data to the household -- 
# Create dataframe to append data into 
pbc_households <- screened_households %>% distinct(hhnr)

# -- 2018 -- 
pbc_households$Confect_P1 <- P1_confectionery$purchases[match(pbc_households$hhnr,
                                                             P1_confectionery$hhnr)]
pbc_households$Confect_P2 <- P2_confectionery$purchases[match(pbc_households$hhnr,
                                                             P2_confectionery$hhnr)]

# -- 2019 -- 
pbc_households$Confect_P3 <- P3_confectionery$purchases[match(pbc_households$hhnr,
                                                             P3_confectionery$hhnr)]
pbc_households$Confect_P4 <- P4_confectionery$purchases[match(pbc_households$hhnr,
                                                             P4_confectionery$hhnr)]
# -- 2020 -- 
pbc_households$Confect_P5 <- P5_confectionery$purchases[match(pbc_households$hhnr,
                                                             P5_confectionery$hhnr)]
pbc_households$Confect_P6 <- P6_confectionery$purchases[match(pbc_households$hhnr,
                                                             P6_confectionery$hhnr)]

# -- 2021 -- 
pbc_households$Confect_P7 <- P7_confectionery$purchases[match(pbc_households$hhnr,
                                                             P7_confectionery$hhnr)]
pbc_households$Confect_P8 <- P8_confectionery$purchases[match(pbc_households$hhnr,
                                                             P8_confectionery$hhnr)]

# -- Code the purchases -- 
# ensure NAs are 0s (this means no sale made)
sum(is.na(pbc_households)) #5141 
pbc_households[is.na(pbc_households)] <- 0 

# split the house numbers away 
pbc_household_split <- select(pbc_households, -hhnr)
hhnrs_pbc <- pbc_households %>% select(hhnr)

# set everything that isn't a 0 to 1 (meaning purchase made)
pbc_household_split[pbc_household_split != 0] <- 1

# bind the columns back together 
pbc_households <- cbind(hhnrs_pbc, pbc_household_split)
```

```{r find_hhnrs_exclude}
# -- Step 5: Check number of households who have made pb_meat purchases -- 
# -- Find households who made no pb_meat purchases and exclude them from data -- 
pbc_households <- pbc_households %>% 
  mutate(Total = select(., Confect_P1:Confect_P8) %>% rowSums(na.rm = TRUE)) 

# -- Exclude households who made no purchases -- 
excluded_confectionery <- pbc_households %>% subset(Total < 1) # 125 households 

# Thus 1034 households meet the confectionery criteria 
```

```{r clean_confectionery}
rm(P1_confectionery, P2_confectionery, P3_confectionery, P4_confectionery, P5_confectionery,
   P6_confectionery, P7_confectionery, P8_confectionery)
```

#### Cleanups 
```{r}
rm(hhnrs_pbc, hhnrs_pbm, hhnrs_pbr)
```

#### Exclude irrelevant households 
We're interested in seeing how households move from the products in one category (e.g., dairy) to make purchases in another (e.g., meat). This helps to identify if their interest is in plant-based products, or if their plant-based sale is perhaps instead due to a dietary preference or allergy. 

To account for this with the data, we'll exclude households which have not made purchases in at least 2 of the plant-based categories: 
```{r identify_exclusion_hhnrs}
# -- Get a list of the hhnrs excluded per category -- 
excluded_dairy <- excluded_dairy %>% select(hhnr)
excluded_meat <- excluded_meat %>% select(hhnr)
excluded_readymeal <- excluded_readymeal %>% select(hhnr)
excluded_confectionery <- excluded_confectionery %>% select(hhnr)

# -- bind them together into one list -- 
excluded_hhs <- rbind(excluded_dairy, excluded_meat, excluded_readymeal, excluded_confectionery) 

# -- Count how many times the household has been excluded -- 
excluded_hhs <- excluded_hhs %>% group_by(hhnr) %>% mutate(count = n())

# -- Remove duplicates so that each row represents a distinct hh -- 
excluded_hhs <- excluded_hhs %>% distinct() # 932 

# -- Cound how many will be excluded due to our criteria -- 
sum(excluded_hhs$count == 4) # 21 households didn't fit the criteria in all categories 
sum(excluded_hhs$count == 3) # 119 households didn't fit the criteria in 3 categories 

# -- Subset the households to exclude -- 
to_exclude <- excluded_hhs %>% subset(count >= 3) # 140 households 

```

```{r filter_data}
# -- Dairy -- 
# Separate the household data to exclude 
dairy_to_exclude <- pb_households[pb_households$hhnr %in% to_exclude$hhnr,]
# Split this from the coded data 
pb_households <- anti_join(pb_households, dairy_to_exclude) # 1019 obs 

# -- Meat -- 
# Separate the household data to exclude 
meat_to_exclude <- pbm_households[pbm_households$hhnr %in% to_exclude$hhnr,] 
# Split this from the coded data
pbm_households <- anti_join(pbm_households, meat_to_exclude) # 1019 obs 

# -- Ready Meals -- 
# Separate the household data to exclude 
readymeals_to_exclude <- pbr_households[pbr_households$hhnr %in% to_exclude$hhnr,] 
# Split this from the coded data
pbr_households <- anti_join(pbr_households, readymeals_to_exclude) # 1019 obs 

# -- Confectionery -- 
# Separate the household data to exclude 
confectionery_to_exclude <- pbc_households[pbc_households$hhnr %in% to_exclude$hhnr,]
pbc_households <- anti_join(pbc_households, confectionery_to_exclude)

```

```{r save_coded_data}
# Split away the "Total" column and save the data 
coded_dairy <- pb_households %>% select(1:9)
coded_meat <- pbm_households %>% select(1:9)
coded_readymeals <- pbr_households %>% select(1:9)
coded_confectionery <- pbc_households %>% select(1:9)

# Rename the columns 
coded_dairy <- rename_columns(coded_dairy)
coded_meat <- rename_columns(coded_meat)
coded_readymeals <- rename_columns(coded_readymeals)
coded_confectionery <- rename_columns(coded_confectionery)

# save into the coded_data folder 
write.csv(coded_dairy, "data/coded_data/dairy.csv", row.names = F)
write.csv(coded_meat, "data/coded_data/meat.csv", row.names = F)
write.csv(coded_readymeals, "data/coded_data/readymeals.csv", row.names = F)
write.csv(coded_confectionery, "data/coded_data/confectionery.csv", row.names = F)

```

The data has now been coded and filtered to be fed into the model. 
